---
title: "Practice Coursework"
author: "Tanheed Rahman"
format:
  html:
    toc: true
    highlight: tango
    self-contained: true
  pdf: default
format-links: false
---

# The palmer analysis

Some analysis on palmer penguins was done on 01/10/25 and it will be shared here.

```{r, echo = FALSE}
#| warning: false
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)

penguins <- palmerpenguins::penguins
```

## Task 1

The first task is to adapt the data frame by adding an additional column indicating if there is any missing data.

The adjusted `penguins` data frame look like this:

```{r, echo = FALSE}
#| label: penguinsTable
#| tbl-cap: The `penguins` data frame

penguins$any_missing_values <- apply(X = penguins, MARGIN = 1, FUN = anyNA)
knitr::kable(penguins[1:6,c(1,3,9)],
             align=rep('c', 3),
             col.names = c("Species", "Bill length (mm)", "Missing values"))
```

## Task 2

The second task is to compute the average bill and flipper length. From this, we get the following data:

```{r, echo = FALSE}
mean_by_species <- aggregate(cbind(bill_length_mm, flipper_length_mm) ~ species,
                             data = penguins,
                             FUN = function(x) mean(x, na.rm = TRUE))
knitr::kable(mean_by_species,
             align=rep('c', 3),
             col.names = c("Species", "Bill length (mm)", "Flipper length (mm)"))
```

Alongside this, we also aim to create the general function `species_mean(df, var)` which calculates the mean of a numeric or integer variable for each species. The code for which, looks like the following:

```{r}
species_mean <- function(df, var){
  stopifnot(class(df[[var]]) %in% c("numeric", "integer"))
  by_spec <- aggregate(df[[var]],
                       by = list(species = df$species),
                       FUN = function(x) mean(x, na.rm = TRUE))
  by_spec
}
spec_mean <- species_mean(penguins, "bill_length_mm")
```

```{r, echo = FALSE}
knitr::kable(spec_mean,
             align=rep('c', 2),
             col.names = c("Species", "Mean bill length (mm)"))
```

## Task 3

The final task is to create a scatter plot of flipper length vs body mass, differentiated by species. In the figures below, the average for each species is represented in black.

```{r, echo = FALSE}
#| warning: false

flipper_means <- aggregate(cbind(flipper_length_mm, body_mass_g) ~ species,
                             data = penguins,
                             FUN = function(x) mean(x, na.rm = TRUE))

graph <- ggplot(penguins, aes(x = flipper_length_mm,
                               y = body_mass_g,
                               color = as.factor(species),
                               shape = as.factor(species))) +
  geom_point(size = 2) +
  labs(x = "Flipper length (mm)", y = "Body mass (g)") +
  labs()
graph + geom_point(flipper_means, mapping = aes(x = flipper_length_mm,  y = body_mass_g),
             color = "black",
             size = 5,
             show.legend = FALSE)
```
