---
title: "Practice Coursework"
author: "Tanheed Rahman"
format:
  html:
    toc: true
    highlight: tango
    self-contained: true
  pdf: default
format-links: false
---

```{r, echo = False}
#| warning: false
library('palmerpenguins')
library('tidyverse')
```

#Using the palmer analysis

Some analysis on palmer penguins was done on 01/10/25 and it will be shared here.

```{r, echo = False}
penguins <- palmerpenguins::penguins
```

##Task 1

The first task is to adapt the data frame by adding an additional column indicating if there is any missing data.

The before and after of the `{r}penguins` data frame look like this:

```{r, echo = False}
penguins
penguins$any_missing_values <- apply(X = penguins, MARGIN = 1, FUN = anyNA)
penguins
```

##Task 2

The second task is to compute the average bill and flipper length. From this, we get the following data:

```{r, echo = False}
mean_by_species <- aggregate(cbind(bill_length_mm, flipper_length_mm) ~ species,
                             data = penguins,
                             FUN = function(x) mean(x, na.rm = TRUE))
mean_by_species
```